An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-21 VLDB2026/main.tex
Update main.tex

	\section{Experiments}
	\subsection{Experimental Setup}
	\subsubsection{Datasets}

	For attribute filtering tasks, we employ 6 real-world datasets widely adopted in existing literature: Msong~\cite{msong2011}, Audio~\cite{audio_unknown}, SIFT1M~\cite{sift2010}, GIST1M~\cite{sift2010}, GloVe~\cite{GloVe2015}, Enron~\cite{enron2015}, covering a diverse range of domains. We independently generate attribute for each dataset.
	For range filtering queries, \textcolor{red}{we evaluate 1 widely used real-world dataset: %Deep~\cite{yandex_deep_dataset}, YT-Audio~\cite{youtube8m_dataset}, and 
	WIT~\cite{wit_dataset}. Following prior work~\cite{DSG}, we use the data point ID (the primary key, ranging from $0$ to $n-1$) within each dataset as its attribute. This ensures that the data is initially sorted.}

	\textcolor{violet}{In addition to the aforementioned datasets, we further evaluated the performance of various algorithms on the Out-of-Distribution (OOD) dataset Text2Image~\cite{texttoimage} and the large-scale dataset Deep100M to validate the algorithm generalization capability. Specifically, the Text2Image dataset comprises two versions: the original version contains 1 million image embeddings as the base dataset and uses 10 thousand text embeddings as the query set; the Text2Image-Mix version is composed of 0.5 million image embeddings and 0.5 million text embeddings as the base dataset, with a query set of 5 thousand image embeddings and 5 thousand text embeddings.}

	\textcolor{violet}{Given that the preceding datasets primarily utilized synthetic attributes, we additionally introduced two datasets containing real attributes for a more comprehensive evaluation: the YT-Audio-Real dataset leverages URLs provided by the YouTube8M dataset to crawl corresponding video category IDs and view counts as its real attributes; whereas for the WIT-Real dataset, we encoded the \texttt{page\_title} field into vectors using the \texttt{clip-vit-base-patch32} model, and used the language type and the length of the \texttt{context\_page\_description} field as real attributes.}

@@

		\label{tab:datasets_combined}
		\resizebox{\columnwidth}{!}{
			\begin{tabular}{cccccc}
				\toprule
				Dataset & Dimension & Base Data & Queries & LID & Type \\
				\midrule
				Msong & 420 & 992,272 & 200 & 23& Audio \\
				Audio & 192 & 53,387 & 200 & 14& Audio \\
				SIFT1M & 128 & 1,000,000 & 10,000 & 19& Image \\
				GIST1M & 960 & 1,000,000 & 1,000 & 45& Image \\
				GloVe & 100 & 1,183,514 & 10,000 & 47& Text \\
				Enron & 1369 & 94,987 & 200 & 23& Text \\
				%Deep & 96 & 1,000,000 & 10,000 & 22& Image \\
				%YT-Audio& 128 & 1,000,000 & 10,000 & 15& Audio \\
				YT-Audio-Real& 128 & 1,000,000 & 10,000 & 15& Audio \\
				WIT & 2048 & 1,000,000 & 40,300 & 48& Image \\
				WIT-Real & 512 & 1,000,000 & 40,300 & \textcolor{blue}{xx}& Image \\
				Text2Image & 200 & 10,000,000 &10,000 & 59& Text,Image \\
				Text2Image-Mix & 200 & 1,000,000 &10,000 & 59& Text,Image \\
				Deep100M & 96 & 100,000,000 & 10,000 & 22& Image \\
				\bottomrule
			\end{tabular}
		}
	\end{table}

@@

		\item \textbf{Recall@k}: The proportion of overlap between the returned approximate $k$ nearest neighbors and the ground-truth $k$ nearest neighbors.
		\item \textbf{QPS (Queries Per Second)}: The number of queries processed per second.
		\item \textbf{Index Construction Time}: The total time required to transform the raw dataset into a queryable index structure.
		\item \textbf{Index Size}: The storage size of the persisted index on disk.
		\item \textbf{Peak Memory Usage}: The maximum memory consumption observed during index construction \textcolor{orange}{and search}.
	\end{enumerate}

@@

	%To investigate the performance of different attribute filtering algorithms during index construction,
	\textcolor{orange}{To explore the time and space overhead of different attribute filtering algorithms during index construction and search,}
	 we evaluate \textcolor{orange}{4} key metrics in the single-attribute scenario: index construction time, index size, \textcolor{orange}{build peak memory and search peak memory}. Notably, PASE on GIST1M and Enron is omitted, as it supports only data with dimensionality less than 512.

@@

	\textit{\textbf{Build peak memory.}}
	As shown in Figure~\ref{fig:build_memory_mb_comparison}, 
	%on small-scale datasets, all methods except NHQ exhibit low memory usage. 
	\textcolor{orange}{NHQ has the highest peak memory usage on almost all datasets.}
	NHQ adopts the kgraph approach for index construction, which requires loading the entire graph data into memory and maintaining a complete graph structure, leading to relatively high memory overhead.

@@

	\textit{\textbf{Different Datasets.}}  
	To evaluate the performance of various methods across different datasets, we conduct experiments on 7 datasets, including one dataset Text2Image for OOD queries. Our analysis focuses on performance variations concerning dataset size, LID, and vector dimensionality.

@@

	%Similar to the evaluation of \textit{Attribute Filtering} algorithms, 
	Similarly, we also evaluate \textcolor{orange}{4} key metrics mentioned above in \textit{Range Filtering} algorithms. \textcolor{orange}{Note: For the YT-Audio dataset with synthetic attributes and the YT-Audio dataset with real attributes, we found that the time and space overhead data of each algorithm are basically the same, which shows that the authenticity of the attribute has little effect on it.
	However, the WIT dataset with synthetic labels is 2048-dimensional, while the WIT-Real dataset with real labels is 512-dimensional, so the dimension will cause a big difference.}

@@

	In this experiment, we follow the query range definition adopted by prior work~\cite{HQI}. \textcolor{red}{Specifically, for a given query, the range ratio is defined as $2^{-i}$	when the query range covers $\frac{n}{2^i}$ data points, where $n$ is the total dataset size.} Based on this definition, we evaluate algorithm performance under 4 query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. 
	%\textcolor{red}{Some algorithms don't support using actual attributes as search keys, instead requiring an ID (the primary key, ranging from $0$ to $n-1$). Therefore, we uniformly use this ID as the search key. This ID is pre-sorted according to its corresponding real attribute, allowing us to map an attribute to its ID without loss of generality for efficient range queries.}

>>>>

	\section{Experiments}
	\subsection{Experimental Setup}
	\subsubsection{Datasets}
	For attribute filtering tasks, we employ 6 real-world datasets widely adopted in existing literature: Msong~\cite{msong2011}, Audio~\cite{audio_unknown}, SIFT1M~\cite{sift2010}, GIST1M~\cite{sift2010}, GloVe~\cite{GloVe2015}, Enron~\cite{enron2015}, covering a diverse range of domains. We independently generate attribute for each dataset.

	\textcolor{violet}{For range filtering queries, we evaluate 1 widely used dataset: WIT~\cite{wit_dataset}. Following prior work~\cite{DSG}, we use the data point ID within each dataset as its attribute. This ensures that the data is initially sorted.}

	\textcolor{violet}{Additionally, we introduced two datasets with real attributes for a more comprehensive evaluation of various attribute filtering and range query algorithms: The YT-Audio-Real dataset leverages URLs provided by the YouTube8M dataset to crawl corresponding video category IDs and view counts as its real attributes; whereas for the WIT-Real dataset, we used the \texttt{clip-vit-base-patch32} model to encode the \texttt{page\_title} field into vectors, and used the language type and the length of the \texttt{context\_page\_description} field as real attributes. It should be noted that, during range filtering, due to the limitations of some algorithms, we uniformly sorted the vectors according to the query attribute.}

	\textcolor{violet}{To validate the algorithm generalization capability, we also further evaluated the performance of various algorithms on the Out-of-Distribution (OOD) dataset Text2Image~\cite{texttoimage} and the large-scale dataset Deep100M. Specifically, the Text2Image dataset comprises two versions: the original version contains 1 million image embeddings as the base dataset and uses 10 thousand text embeddings as the query set; the Text2Image-Mix version is composed of 0.5 million image embeddings and 0.5 million text embeddings as the base dataset, with a query set of 5 thousand image embeddings and 5 thousand text embeddings.}

@@

		\label{tab:datasets_combined}
		\resizebox{\columnwidth}{!}{
			\begin{tabular}{cccccc}
				\toprule
				Dataset & Dimension & Base Data & Queries & LID & Type \\
				\midrule
				Msong & 420 & 992,272 & 200 & 23& Audio \\
				Audio & 192 & 53,387 & 200 & 14& Audio \\
				SIFT1M & 128 & 1,000,000 & 10,000 & 19& Image \\
				GIST1M & 960 & 1,000,000 & 1,000 & 45& Image \\
				GloVe & 100 & 1,183,514 & 10,000 & 47& Text \\
				Enron & 1369 & 94,987 & 200 & 23& Text \\
				%Deep & 96 & 1,000,000 & 10,000 & 22& Image \\
				%YT-Audio& 128 & 1,000,000 & 10,000 & 15& Audio \\
				YT-Audio-Real& 128 & 1,000,000 & 10,000 & 15& Audio \\
				WIT & 2048 & 1,000,000 & 40,300 & 48& Image \\
				WIT-Real & 512 & 1,000,000 & 40,300 & 48& Image \\
				Text2Image & 200 & 10,000,000 &10,000 & 59& Text,Image \\
				Text2Image-Mix & 200 & 1,000,000 &10,000 & 59& Text,Image \\
				Deep100M & 96 & 100,000,000 & 10,000 & 22& Image \\
				\bottomrule
			\end{tabular}
		}
	\end{table}

@@

		\item \textbf{Recall@k}: The proportion of overlap between the returned approximate $k$ nearest neighbors and the ground-truth $k$ nearest neighbors.
		\item \textbf{QPS (Queries Per Second)}: The number of queries processed per second.
		\item \textbf{Index Construction Time}: The total time required to transform the raw dataset into a queryable index structure.
		\item \textbf{Index Size}: The storage size of the persisted index on disk.
		\item \textbf{Peak Memory Usage}: The maximum memory consumption observed during index construction.
	\end{enumerate}

@@

	To investigate the performance of different attribute filtering algorithms during index construction, we evaluate \textcolor{orange}{4} key metrics in the single-attribute scenario: index construction time, index size, \textcolor{orange}{build peak memory and search peak memory}. Notably, PASE on GIST1M and Enron is omitted, as it supports only data with dimensionality less than 512.

@@

	\textit{\textbf{Build peak memory.}}
	As shown in Figure~\ref{fig:build_memory_mb_comparison}, 
	%on small-scale datasets, all methods except NHQ exhibit low memory usage. 
	\textcolor{orange}{NHQ has the highest peak memory usage on almost all datasets.}
	This is because NHQ adopts the kgraph approach for index construction, which requires loading the entire graph data into memory and maintaining a complete graph structure, leading to relatively high memory overhead.

@@

	\sout{\textit{\textbf{Different Datasets.}}  }
	To evaluate the performance of various methods across different datasets, we conduct experiments on 7 datasets, including one dataset Text2Image for OOD queries. Our analysis focuses on performance variations concerning dataset size, LID, and vector dimensionality.

@@

	%Similar to the evaluation of \textit{Attribute Filtering} algorithms, 
	Similarly, we also evaluate \textcolor{orange}{4} key metrics mentioned above in \textit{Range Filtering} algorithms. \textcolor{orange}{For the YT-Audio dataset with synthetic attributes and the YT-Audio dataset with real attributes, we can found that the data for index construction is basically the same, indicating that the authenticity of the attribute has almost no effect on index construction.
	However, the WIT dataset with synthetic labels is 2048-dimensional, while the WIT-Real dataset with real labels is 512-dimensional, so the dimension will cause a big difference.}

@@

	In this experiment, we follow the query range definition adopted by prior work~\cite{HQI}. \textcolor{red}{Specifically, for a given query, the range ratio is defined as $2^{-i}$	when the query range covers $\frac{n}{2^i}$ data points, where $n$ is the total dataset size.} Based on this definition, we evaluate algorithm performance under 4 query range settings: $2^{-2}$, $2^{-4}$, $2^{-6}$, and $2^{-8}$. \textcolor{red}{Some algorithms don't support using actual attributes as search keys, instead requiring an ID (the primary key, ranging from $0$ to $n-1$). Therefore, we uniformly use this ID as the search key. This ID is pre-sorted according to its corresponding real attribute, allowing us to map an attribute to its ID without loss of generality for efficient range queries.}
